---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

1. Generating random variables: \newline
   p{distr} generates a cdf; 
   d{distr} generates a pdf/pmf
   q{distr} generates quantiles
   r{distr} generates a sample
   
   {distr} can be any distribution:
    - norm (normal)
    - exp  (exponential)
    - unif (uniform)
    - gamma (gamma)
    - chisq   (chi-square)
    - binom (binomial)
    - geom  (geometric)
    - pois  (Poisson)
   
Example   
   
```{r}
# n is the size of a sample
n <- 10
# m is no. of repetitions
m <- 1000
# generating die throws
diethrows <- sample(1:6, prob = c(.1,.2,.2,.3,.1,.1), size = n, replace = T); diethrows
# means of those; dtm ="diethrowmeans"
dtm <- replicate(m, mean(sample(1:6, prob = c(.1,.2,.2,.3,.1,.1), size = n, replace = T)))
#expected value
expval <- sum((1:6)*c(.1,.2,.2,.3,.1,.1)); expval
# summary of dtm
summary(dtm)
# histogram
hist(dtm,
     probability = T,
     breaks = 20,
     main = paste("Means of n =", n, "throws of a biased die"),
     col = "lightblue",
     xlab = "Means",
     xlim = c(min(dtm),max(dtm))
     )
# superimpose normal distribution
x <- seq(min(dtm),max(dtm), by = .01)
lines(x, dnorm(x, mean = mean(dtm), sd = sd(dtm)), col ="red", lwd = 2)
# does it agree with the LLN?
sum(abs(dtm - expval)>.1)/m
```


2 Task 2

1. Generate a sample of size 1000 combined from two samples of size 500 from the normal distributions N(mu,5) with mu = 0 and mu = 1 
2. Calculate sample mean, sample standard deviation and compare with the exact values
3.
4. For q= 0.1,...,0.9 calculate sample q-percentiles manually 
  (i.e. by sorting the data and picking the corresponding values) and by using percentile
5. Compare with theoretical quantiles of that distribution
6. Check the k-sigma rule for k=1,2,3
7. Draw a histogram of the data
8. Draw the theoretical density 

```{r}
#   Your code goes here
```

