---
title: "Hypothesis testing for parameters of normal distribution"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

**Example 1** Testing
$$
  H_0\,: \mu = \mu_0 (=0) \quad \textrm{vs} \quad H_1\,: \mu = \mu_1 (= 1)
$$
for normal distribution $\mathscr{N}(\mu,\sigma^2)$ with **known** $\sigma^2$  

This is a $z$-test, based on the fact that under $H_0$ the distribution of the $Z$-statistics is standard normal:
$$
    Z : = \frac{\sqrt{\sigma}}{n}(\overline{\mathbf{X}} - \mu_0) \sim \mathscr{N}(0,1) \qquad {\text{under}} \quad H_0
$$
Alternative $H_1$ can be *one-sided* ($\mu>\mu_0$ or $\mu < \mu_0$) or *two-sided* ($\mu \ne \mu_0$); the corresponding $z$-test will also be one- or two-sided


```{r}
N <- 100  # sample size
##  data generation; observe that actually two different normal distribution are mixed up
x <- rnorm(N, mean = c(-2,2), sd = c(1,2))
##  IMPORTANT: requires BSDA-package
z.test(x, mu = 0, alternative = "greater", sigma.x = 1.5)

```

**Remark** z-test is rarely used in practice since usually $\sigma$ is not known 

**Example 2** _When $\sigma^2$ is not known, the `t.test` is used; the corresponding $T$-statistics_

$$
  T(X) := \frac{\sqrt{n}}{S}(\overline{\mathbf{X}} - \mu_0) = \frac{Z}{S/\sigma}
$$
has the *Student* $\mathscr{T}_{n-1}$-distribution under $H_0$. Therefore, it is enough to replace $z$-statistics and quantiles with $t$-statistics and quantiles
```{r}
t.test(x, alternative = "two.sided")    ## test the (default) H_0: mu = 0 vs H_1: mu \ne 0
```

**Example 3** _Comparison of two means_

```{r}
# test the hypothesis that mu_x is larger than mu_y
z.test(x=x[x<0], y=x[x>0], sigma.x = 1, sigma.y = 2, alternative = "l")
# clearly, we should stick to the alternative here
```


**Example 4** _Hypothesis testing for variances (one sample)_
$$
  H_0\,: \sigma^2 = \sigma^2_0 \quad \textrm{vs} \quad H_1\,: \sigma^2 > \sigma^2_0 \qquad  (\textrm{or}\  <,\ \ne)
$$

```{r}
## 
## requires package EnvStats
##
varTest(x, sigma.squared = 1, alternative = "g")
```

**Example 5** _Hypothesis testing for variances (two means)_ 

```{r}
var.test(x[x<0],x[x>0], alternative = "l")
```


**Problem 1** Hypothesis testing for normal distribution $\mathscr{N}(\mu,\sigma^2)$   
1. Generate $100$ values of data `x` with exponential distribution $\mathscr{E}(\lambda)$  for  
- (a) $\lambda=1$   
- (b) $\lambda = 5$  
2. Form `y` as `-2*x`    
3. Test hypotheses  

  (a) $H_0\,: \mu = 1 \quad \textrm{vs} \quad H_1\,: \mu < 1$  
  (b) $H_0\,: \mu_x = \mu_y \quad \textrm{vs} \quad H_1\,: \mu_x > \mu_y$  
  (c) $H_0\,: \sigma^2 = 1 \quad \textrm{vs} \quad H_1\,: \sigma^2 > 1$  
  (d) $H_0\,: \sigma^2_x = \sigma_y^2 \quad \textrm{vs} \quad H_1\,: \sigma^2_x < \sigma_y^2$ 


4. For each test, provide a discussion of the results (p-value etc) and the decision to be made  
   Comment on whether or not the results are expected in view of the special way the data are formed
   
5. Submit results as an `R Notebook`

```{r}
## Your code and explanations here
```


